Speculation in foreign exchange: noise or information

### 2015-09-08 Knitr/NewPaper.Rnw
Extreme2.R is the latest on extremes This needs to be developed so that there are distributions. The paper has been re-structured.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study is used to assess the views of Keynes and Friedman on speculation. Speculative extremes are ranked by intensity of sentiment and weight of speculative activity. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.  The extremes say little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity. 
\end{abstract}
\section*{Introduction}
There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of this spectrum is the view of  \citet[p. 101]{Keynes1936} in Chapter 12 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''. Here speculation is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that does not emerge randomly, but feeds off itself in an evolution of collective sentiment that encourages trends in asset prices.

Towards the other end of the spectrum, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedman argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.

The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing  the strength of speculative opinion or the weight of speculative positions does not improve forecasts of future returns. This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence seems to help rather than hinder the incorporation of information into prices.

\section{Literature}
\subsection*{Overview}
  Speculation is an important concept in economics, particular in the area of finance.  However, there is no clear, unambiguous theory of speculation. Speculation plays a key role in three important and related areas: market-microstructure, the rationale for the trading of financial securities and rational expectations.

One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have information. The \citet{Kyle1985Continuous} model has market-makers receiving orders from the informed and the uninformed.  On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.  The bid-ask spread is the payment that the market-maker receives in return for taking the risk of meeting an informed trader.  Ultimately, the uninformed traders make payments to the informed traders with the market-makers taking a cut for the risk and the provision of exchange.  See also \citet{glosten1985bid}.

In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.  There may be institutional and psychological reasons for the lack of attention to fundamental information. Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value. This may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy. Noise is more psychological.

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninformed speculation provides a rational and an opportunity for the trading of financial securities. There must be different opinions.  Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, prepared to take the other side of the transaction, the informed can take advantage of their information.  This creates a justification for acquiring information. Speculative noise-traders can therefore create a gap between the price and the fundamental value that will encourage informed traders to pay the cost of acquiring information.   \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value. This begs the question: are speculators trading on noise or are they informed?

  Speculation is also an important part of the argument for why prices should tend towards a level that consistent with \emph{the rational expectations equilibrium} (RER). This is the level that is consistent with rational expectations.  In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent. These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the mechanism through which equilibrium is achieved.

However, though equating noise and noise-trading as a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, it may also be an over-simplification that is inappropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.

  The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.  Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information into the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than to form an independent view of the actual underlying, fundamental value \citet[p.101]{Keynes1936}.

\subsection{Empirical work}
Empirical assessment of speculation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish those that may be noise-traders from those assumed to be informed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  For example speculation can be identified with momentum trading. It is also possible to assume that certain investors, such as investment banks and hedge funds, are informed while other trade without information. The latter group may include day-traders or private individuals.  
It has also been suggested that speculative techniques can change over time.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation can be be associated with sentiment and measures of sentiment are then used as ways of measuring speculation.

%Noise stragegies may be successful for a while.  Momentum trades tend to be successful for certain ties.  Also discussion the literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.  These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders.

\subsection{Why this is important}
The Keynesian view of speculation is that it is an addition to market noise that makes it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. As such, it exacerbates the tendency towards putting too much emphasis on the immediate and works against longer-term and strategic thinking.  This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).  There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.  For example, regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they limit risk-taking and encourage more long-term attitudes towards investment.  However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.

The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.  The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.  The argument for a tax on financial transactions depends on the idea that speculators will drive the price away from its fundamental value.

%volcker rule?

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the foreign exchange market.  The analysis will involve three steps:  assessment of the price implication of the noise-trader model;  measurement of speculative activity in the foreign exchange market; finally, assessment of whether extreme speculative market activity is consistent with the underlying model that is used.  If the speculative activity in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is closer to the Keynesian view than that of Friedman.

The \citet{Delong1990noise} noise-trader model used in this study. This is a derivative of the \citet{Samuelson1958} overlapping generations model with two-period agents making a portfolio choice between two assets: a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; a risky asset $(u)$ which also pays the same dividend but is not in elastic supply.  There are two types of investor: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

\begin{equation}
\rho_t \sim N(\rho^*, \sigma_{\rho^2})
\end{equation}

Maximising utility with respect to holdings of the sage and risky assets for the informed and uninformed traders and iterating backward to solve for the current value,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma}{r}\ \big (_t\sigma^2_{p,t+1} \big )
\label{eq:NP1}
\end{equation}

The deviation from the true fundamental value of unity is a result of the discounted value of the noise-traders' misperception of the price $(p_t - \rho^*)$, multiplied by the weight of noise-traders in the total $(\mu)$ plus a perpetuity of the average noise-trader misperception $(\frac{ \mu \rho^*}{r} )$, adjusted for the risk.   This risk is a function of the volatility of the price which,  from Equation \ref{eq:NP1} is a consequence of the variability of the misperception of speculative noise-traders.   All but the second term in equation \ref{eq:NP1} are constants, and if the variance of the price one period ahead  is equal to the variance of the misperception,

\begin{equation} 
 _t\sigma^2_{p,t+1} = \frac{\mu^2 \sigma_{\rho}^2}{1+r^2}  
\end{equation}

the final price equation becomes:

\begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
\label{eq:NP2}
\end{equation}

The following four effects are noted by \citet[pp. 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} is the \emph{noise-trader risk} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information. In this model, the more numerous the noise-traders, the more volatile prices.

The Friedman-effect and the Price-pressure are the effects that are studied here.  Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversal.  The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.  Create-space effect would suggest that when the weight of noise-traders is high and extreme, there is greater incentive and opportunity for the informed to take advantage of the divergence between fundamental value and price.

\subsection{The implications of the noise-trader model}
There are two parameters from the noise-trader price equation \ref{eq:NP2} that can be used to measure the amount of speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  The aim of this section is to  identify the returns that are implied by the noise-trader model when either the intensity of the misperception or the weight of noise-traders is increased to extreme levels so that this can be used as a benchmark against which foreign exchange market activity can be judged.

To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of misperception-perceptions and noise-trader weights.  Two time series are created.  The first series is based on a random noise-trade misperception, $(\rho^*)$.  This is called the Misperception Model (MPM). The second is based on a random weight of noise-traders $(\mu)$ and is called the Random Weight Model (RWM).   The simulation creates a 1000 period price series from Equation \ref{eq:NP2}. The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citet[p. 127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader misconception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.   
% it is possible to have other specifications. For example, the noise-traders oculd become more prominent if their strategy is more successful,

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when misperception-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study. Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

\subsection{The results of simulation}
The results of the simulation are recorded in Table \ref{tabref:MC1}.  It shows that the mean return to an investment strategy in the whole of the sample is very close to zero.  However, for each of the extreme readings, there are large negative returns to follow. This is consistent with the structure of the model: noise-traders drive price above intrinsic value and subsequent returns will be negative; the more extreme the misperception or the weight of uninformed speculative noise-traders in the market, the greater the negative return that is likely to be recorded.

\begin{table}[h!]
\begin{threeparttable}
\caption{Results of the Monte Carlo Simulation of the Noise-Trader Model}
\begin{tabular}{p{4cm} | r r r r}
 \label{tabref:MC1}
                            & Base Case  & Extreme (MPM) & Extreme (RWM) \\
  \hline
\hline
Mean                     & -0.00 & -0.90*   & -0.49 \\
Standard Deviation  & 0.98 &  0.49*     &  0.42*\\
Skew                      &  -0.07 &  -0.16   &  -0.78*\\
Kurtosis                  &  -0.14 &  -0.38    &  -0.35\\
Number                 &  1000        &   50        & 50\\

>>>>

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
  If speculation is just noise, price movements associated with extremes of speuclative sentiment or heavy involvement of speculative activity should be reveversed; if speculation is informed, associated price movement is more likely to be sustained.  A unique set of risk-reversals on currency options is used to measure speculative sentiment and regulatory positions measure speculative activity. An event study shows that extremes do not increase the probability of a price reversal, supoprting the view that speculation is part of the price disovery process.     
\end{abstract}
\section*{Introduction}
Ideas about the way that speculation affects financial market span a broad range.  Towards one end is the Keynesian view that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment; It is noise that emerges from the interaction of collective sentiment to encourages trends in asset prices.

Towards the other end of the spectrum, Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.

This paper contributes to the debate about the nature of speculation in a number of ways. A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speulative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.  Speculative extrmes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be chracterised as a random walk; this suggests that speculation is part of the process of price discovery.

For example, Figure \ref{fig:ES1} show the evolution of a selection of foreign exchange prices after an exrtreme reading of speculative sentiment. Sentiment is measured by an extreme in the price of currency call options relative to the equivalent put.  It is clear that the foreign exchange rate has a contemporaneous movement with the sentiment but that after reading the extreme, there is no clear direction.  The lack of clear direction indicates lack of information.  As such, extreme speuclative positions do presage reversales, indicating that they are not associated with divergence from equilibrium.

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.6]{RRCum16}
\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the hghest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.  The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.  If speculation is uninformed, extremes shoudl be followed by reversals; if specuation is informed, extremes are more likely to be followed by a random walk.}
\label{fig:ES1}
\end{figure}
% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

The nature of speculation is central to many contemporary questions about international financial markets. For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.  If speculation is largely uninformed it will add to the level of noise and maker it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. It will increase the tendency towards activity that is focused on the short-term.  However, if speculation is informed and part of the process of price discovery, regulatory measures may reduce liquidity and undermine the pricess of price discovery. Increased volatilty, heightened risk and reduced transparency together with other costs may provide an countervailing force.

The rest of the paper proceeds as follows:  Section Two analyses the distinction between informed and uninformed speuclators; Section Three discusses the measurement of speculation and the testing of extreme positions; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.

\section{Informed and uninformed traders}
Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities and rational expectations. However, there is no clear, unambiguous theory of speculation.  While speculation may be viwed as being informed or uninformed, those trading without information are often categorised as being either \emph{noise-traders} or \emph{liquidity traders}.

Noise-trading is associated with psychogical biases and restricted information while liquidity-trading is about institutional constraints.  Therefore, noise is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools. See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{represetntativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.  Market-participants do not make decisions independently of those made by others. Information is inferred from the activities of others. \citet[p.101]{Keynes1936} said investment decisions could be based on the expecttions of others' valuations.  institutional constraints may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy. The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collaeral constraints can force investment decisions.  However, the terms are not fixed and  \citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.

An important part of microstructure research is the distinction between informed and uninformed traders.  The focus of the \citet{Kyle1985Continuous} model is a market-maker receiving orders from the informed and the uninformed.  On the assumption that uniformed orders arrive in a random fashion, the weight of orders implies something about the knowledge of the informed and should led to adjustment of the bid-ask spread as part of the process of price discovery.  This bid-ask spread is the payment that the market-maker receives from the uninformed in return for taking the risk of meeting an informed trader.  See also \citet{glosten1985bid}.

As such, Uninformed activity provides a rationale and the opportunity for trading. Unless there is a difference of opinion about value, there is no reason to trade. Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.  Indeed, this space can encourage informed traders to pay the cost of acquiring information.  See, for example \citet{Grossman1980Impossibility}.

\citet[p.529]{BlackNoise} suggested that this interaction of informed and uninformed activity was essential to the smooth functioning of financial markets. Noise, he argued, was the  ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

\citet{EasleyPIN} use informaton as an explanation for the effect of large orders on stock prices.  The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker recieves a large order; it may also be a function of the market-maker knowledge that large orders are more likely to informed traders seeking to take advantage of their informational advantage.  \citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.  The find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that this is more than offset by the high quantity of uninformed traders.  Low liquidity stock, by contrast, have fewer information events but informed traders have a much larger effect on prices.

Informed speculation is also used as the theoetical justification for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.  Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate or sell forecasts that will tend to drive prices towards the equilibrium \citet{muth1961Rational}.  This tends to leave uninformed speculation as a pure noise that can be characterlised as white-noise random variable. While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.

However, a large proportion of uninformed traders in the market can undermine this process.  \citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where \emph{noise-trader risk} can prevent the arbitrate of informed speculators from driving prices back towards the fundamental value.  The risk that noise-traders will drive prices even further away from equilibrium before the convegence can limit the activity of informed speculators when they are risk averse.\footnote{This three-period, overlapping generation model with informed and uininformed traders has a price equation \begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
\label{eq:NP2}
\end{equation} 
 where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho$ is the price mis-perception of the uninformed traders; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}

The De-Long noise-trader model also provides a useful benchmark against which to assess the activity of speculators. The following four effects are noted by \citet[pp. 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} is the \emph{noise-trader risk} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information. In this model, the more numerous the noise-traders, the more volatile prices.

The Friedman-effect is the activity of informed speulators.  They take advantage of the space that has been created by the price pressure of the uninformed traders to profit from this mis-pricing.  Uniformed traders tend to hold more when their opinions are strong.  This is the noise-trader risk that can prevent informed traders from fulfuillig their arbitrage role. If uninformed traders are very prominent or have strong opinions, prices should show the greatest divergence from fundamental value and the risk of a reversal shoould be at its greatest.

\subsection{Empirical work}
Empirical assessment of speculation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish between the informed and uninformed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  Common strategies that have been used are momentum (ref), contrarian (ref) and carry (ref).   It is also possible to assume that certain investors usually have more information than others.  For example, hedge funds, investment banks and other professionals can be assuemd to be informed while retail or day traders may be uninformed.

%It has also been suggested that speculative techniques can change over time.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation can be be associated with sentiment and measures of sentiment are then used as ways of measuring speculation.

%Noise stragegies may be successful for a while.  Momentum trades tend to be successful for certain ties.  Also discssion the literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.  These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders. 
\citet{EasleyPIN2} propose a model to try to measure informed market activity based on their assertion that the bid-ask spread is a function of the level of informed trading.  As informed traders are most likely to want to take advantage of their informational advantage, large orders are assumed to be informed.\footnote{With market-makers adopting Bayesian reasoning to update their fundamental valuation by using order flow to estimate the probability that there has been an information event and if there has whether it is good news or bad news, the authors seek to operationalise the following model:
\begin{subequations}
\begin{equation}
b(t) = E[V_i|t] - \frac{\mu P_b(t)}{\varepsilon + \mu P_b(t)}(E[V-t|-\underline{V}])
\end{equation}
\begin{equation}
a(t) = E[V_i|t] + \frac{\mu P_g(t)}{\varepsilon + \mu P_g(t)}(\overline{V}E+ [V-t|])
\end{equation}
\end{subequations}
  where $b(t)$ is the bid price at time $t$; $V$ is the value of the security; $\mu$ is the share of informed traders in the market; $P_g(t)$ is the probability that there has been some good news and $P_b(t)$ is the probability that there has been some bad news; $\varepsilon$ is the share of uninformed traders in the market; $\underline{V}$ is the value if there is bad news and $\overline{V}$ is the value if there is good news.}

However, identifying large orders and making an assessment of the level of informed trading based on that is much harder to achieve in an over-the-counter market like foreign exchange. While electronic order books have become more widely used in recent years, much of the activity still takes place through bilateral agreements.  Information about the nature of transactions including size and origin of orders is not widely available\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.  They find that $\$$ 1 billion of net US dollar purchase moves the deutsche mark price of the US unit by 0.5 percent. They argue that the orders cystalise a range of information that is used to value currencies.}    \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.  Their tick test has four classifications: the uptic, the downtick, the zero-uptick and the zero-downtick. A trade is an uptick if the price is higher than the previous trade.  If the price is the same as the previous trade, if the previous trade was classified as an uptick, this trade is a zero-uptick.  
\section{Methodology}
The aim of this study is to assess whether speculation in the foreign exchange market is informed or uninformed. Speculation must be identified and then extreme speculation will then be compared with what would be expected if it were informed or uninformed.  If informed and part of the price-discovery process, extreme speculation would be a function of perceived divergence from equilibrium or new information.  The following price performance should be a random walk.  If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticiapted.

Returning to the \citet[pp. 14-15]{Delong1990noise} model summarised by Equation \ref{eq:NP2}, there are two parameters that can be used to measure speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  If either of these reach an extreme, prices should be expected to reverse\footnote{See \citet{Hayward2013} for full details of simulations of this model.}

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction. In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.%\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.  However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.   The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.  The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.  In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons. Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.  Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citet gc{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
